---
title: "Lecture2"
author: "Me"
date: "01/16/2022"
output:
  pdf_document: default
  html_document: default
---

# heading
## subheading
### subsubheading

# the first chunk in the file
```{r setup, include=FALSE}   
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## R Markdown
**Markdown** is very useful!!! When you write `README.md` files for codes in github repositories, you also need Markdown. It's not limited to R programming!

double hash = heading

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:
```{r cars}
summary(cars)
```

### shortcut keys(very useful!!!)
add a new code chunk: command+option+i

run current chunk: shift+command+enter

comment a line: command+shift+c

## Including Plots
You can also embed plots, for example:

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R codes that generated the plot.
"pressure" is the title of the chunk
```{r pressure, echo=FALSE}
plot(pressure)
```
The codes will be run and the out put will be shown together with the code blocks
```{r}
print("a")
```
## start about facet
```{r}
# draw each panel depending on its class.
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  facet_wrap(~class,nrow=2)  # draw all panels in two rows

```

```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  facet_grid(drv~cyl) # show all combinations of the two variables
```

```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  facet_grid(.~cyl) #fomular  # if you put a "." then you could draw panels with only one variable
```

```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  facet_grid(drv~cyl) # show all combinations of two variables
```

# more geometries
```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy)) 
```

```{r}
ggplot(data=mpg)+
  geom_smooth(mapping = aes(x=displ,y=hwy))   # the default method is "Loess"
```

use "Help" on the right to check all parameters of geom_smooth
```{r}
ggplot(data=mpg)+
  geom_smooth(mapping = aes(x=displ,y=hwy),method="lm")  # if you want a straight line instead of a curve
```

```{r}
ggplot(data=mpg)+
  geom_smooth(mapping = aes(x=displ,y=hwy,linetype=drv))  # now the data are broken into 3 categories based on the type of "drv"
```

```{r}
# multiple layers
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,color=drv))+
  geom_smooth(mapping = aes(x=displ,y=hwy,linetype=drv,color=drv)) 
```

```{r}
# some part appear twice in the codes and they are the same, so we could put them in the first () and you don't need to type them twice
ggplot(data=mpg,mapping = aes(x=displ,y=hwy,color=drv))+ # mapping here applies for everything below
  geom_point()+
  geom_smooth(mapping = aes(linetype=drv)) 
```

```{r}
ggplot(data=mpg,mapping = aes(x=displ,y=hwy,color=drv,linetype=drv))+
  geom_point()+
  geom_smooth() 
```

```{r}
ggplot(data=mpg,mapping = aes(x=displ,y=hwy,linetype=drv))+
  geom_point(mapping = aes(color=class))+
  geom_smooth() 
```

```{r}
# filter the data and draw fitting curve for only one class
ggplot(data=mpg,mapping = aes(x=displ,y=hwy))+
  geom_point(mapping = aes(color=class))+
  geom_smooth(data=filter(mpg,class=="subcompact"),se=FALSE)  # se is the error bar
```

# Bar charts and statistical transformations
```{r}
summary(diamonds)
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut))
```

```{r}
demo<-tribble(
  ~cut,~freq,
  "Fair",1610,
  "Good",4906,
  "Very Good",12082,
  "Premium",13791,
  "Ideal",21551
)
ggplot(data=demo)+
  geom_bar(mapping=aes(x=cut,y=freq),stat="identity") # when stat is "count", then it will count the frequency of data in each group. If it's "identity". it will plot the value of y directly but not count the freq
```
```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity)) # show the parts of the bar according to "clarity"
```

```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),position="fill") # position is how the bars are arranged
```

```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),position="dodge") # dodge means bars avoiding overlapping with each other
```

```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy),position="jitter") # will not look like dots are on the grids # add a fluctuation to x and y
```

```{r}
ggplot(data=mpg)+
  geom_jitter(mapping=aes(x=displ,y=hwy)) # will not look like dots are on the grids
```

