---
title: "Lecture 9"
author: "Me"
date: "3/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# assumptions
- Independence, every observation is independent from others
- Normal/Gaussian distribution of residuals
- Constant variance of residuals -- homoskedasticity    Var is the same for all samples

```{r}
View(PlantGrowth)
```

# t-test
comparing only two groups at a time

```{r}
ggplot(PlantGrowth,aes(group,weight))+
  geom_boxplot(notch=TRUE) #ugly # don't have enough data to do the notching
```
```{r}
ctrls<-PlantGrowth%>%filter(group=="ctrl")
qqnorm(ctrls$weight)
qqline(ctrls$weight) # check whether the quantiles look like a normal distribution
# the theoretical quantiles is a Gaussian distribution
```
```{r}
result<-t.test(ctrls$weight,mu=0) # althouth mean of this dataset doesn't look like 0
result  # at least we reject the NULL hypothesis
```
```{r}
trt1<-PlantGrowth%>%filter(group=="trt1")
result<-t.test(ctrls$weight,trt1$weight)  # p value 0.25, you can't reject the hypothesis that they have same mean
result
```

What if it's not Gaussian distribution?
What if exponantial distributions?
# nonparametric variant: Wilcoxon test
```{r}
result<-wilcox.test(ctrls$weight,mu=0) # so we kown that mean is not 0, but p value is larger this time
result
```
```{r}
trt1<-PlantGrowth%>%filter(group=="trt1")
result<-wilcox.test(ctrls$weight,trt1$weight) 
result
```


# ANOVAs
when you have more than 2 samples and you want to test some variables

```{r}
View(iris)
```

```{r}
ggplot(iris,aes(Species,Sepal.Length))+geom_boxplot(notch = TRUE)
```
```{r}
result<-aov(Sepal.Length~Species,iris) # ANOVA tests
plot(result) # the first two plots are more useful

```
```{r}
summary(result)  # it's not a good idea to type "result" directly and in that case you will miss some info
# result
```

# post hoc test
Tukey's test, Tukey's Honest Significant Difference test
Only when you have the result for ANOVA test can you do Tukey's test.
Compare the samples two by two
```{r}
TukeyHSD(result) # the results show that the three groups are totaly different
```

# bad data
```{r}
result<-aov(breaks~tension,warpbreaks)
plot(result)
```
```{r}
my_warpbreaks<-warpbreaks%>%mutate(log_breaks=log(breaks))  # transform the data
result<-aov(log_breaks~tension,my_warpbreaks)
plot(result)
```
```{r}
summary(result)
```

# nonparametric 
```{r}
result<-kruskal.test(breaks~tension,warpbreaks)
result
```

#Linear mixture model
some random effect, some fixed effect
```{r}
library(lme4)
```

```{r}
str(sleepstudy)
sleepstudy
```

```{r}
ggplot(sleepstudy,aes(Days,Reaction))+
  geom_point()+
  geom_smooth(method="lm")+
  facet_wrap(~Subject)  # draw panels for every factor in "Subject"
```
```{r}
fm1<-lmer(Reaction~Days+(Days|Subject),sleepstudy,REML=FALSE)
plot(fm1)
```
```{r}
summary(fm1)
```
```{r}
fm0<-lmer(Reaction~(Days|Subject),sleepstudy,REML=FALSE)
summary(fm0)
```

```{r}
anova(fm1,fm0)
```








