---
title: "Lecture2"
author: "Me"
date: "2/4/2021"
output: html_document
---

# heading
## subheading
### subsubheading

# the first chunk in the file
```{r setup, include=FALSE}   
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## R Markdown
**Markdown** is very useful!!! When you write `README.md` files for codes in github repositories, you also need Markdown. It's not limited to R programming!

double hash = heading

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:
```{r cars}
summary(cars)
```

### shortcut keys(very useful!!!)
add a new code chunk: command+option+i

run current chunk: shift+command+enter

## Including Plots
You can also embed plots, for example:

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R codes that generated the plot.
"pressure" is the title of the chunk
```{r pressure, echo=FALSE}
plot(pressure)
```
The codes will be run and the out put will be shown together with the code blocks
```{r}
print("a")
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  facet_wrap(~class,nrow=2)
```

```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  facet_grid(drv~cyl) # show all combinations of two variables
```

```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  facet_grid(.~cyl) #fomular
```

```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy))+
  facet_grid(drv~cyl) # show all combinations of two variables
```

```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy)) 
```

```{r}
ggplot(data=mpg)+
  geom_smooth(mapping = aes(x=displ,y=hwy)) 
```

use "Help" on the right to check all parameters of geom_smooth
```{r}
ggplot(data=mpg)+
  geom_smooth(mapping = aes(x=displ,y=hwy),method="lm") 
```

```{r}
ggplot(data=mpg)+
  geom_smooth(mapping = aes(x=displ,y=hwy,linetype=drv)) 
```

```{r}
ggplot(data=mpg)+
  geom_point(mapping = aes(x=displ,y=hwy,color=drv))+
  geom_smooth(mapping = aes(x=displ,y=hwy,linetype=drv,color=drv)) 
```

```{r}
ggplot(data=mpg,mapping = aes(x=displ,y=hwy,color=drv))+
  geom_point()+
  geom_smooth(mapping = aes(linetype=drv)) 
```

```{r}
ggplot(data=mpg,mapping = aes(x=displ,y=hwy,color=drv,linetype=drv))+
  geom_point()+
  geom_smooth() 
```

```{r}
ggplot(data=mpg,mapping = aes(x=displ,y=hwy,linetype=drv))+
  geom_point(mapping = aes(color=class))+
  geom_smooth() 
```

```{r}
ggplot(data=mpg,mapping = aes(x=displ,y=hwy))+
  geom_point(mapping = aes(color=class))+
  geom_smooth(data=filter(mpg,class=="subcompact"),se=FALSE)  # se is the error bar
```

# Bar charts and statistical transformations
```{r}
summary(diamonds)
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut))
```
```{r}
demo<-tribble(
  ~cut,~freq,
  "Fair",1610,
  "Good",4906,
  "Very Good",12082,
  "Premium",13791,
  "Ideal",21551
)
ggplot(data=demo)+
  geom_bar(mapping=aes(x=cut,y=freq),stat="identity")
```
```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity))
```

```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),position="fill")
```
```{r}
ggplot(data=diamonds)+
  geom_bar(mapping=aes(x=cut,fill=clarity),position="dodge")
```
```{r}
ggplot(data=mpg)+
  geom_point(mapping=aes(x=displ,y=hwy),position="jitter") # will not look like dots are on the grids
```

```{r}
ggplot(data=mpg)+
  geom_jitter(mapping=aes(x=displ,y=hwy)) # will not look like dots are on the grids
```

