---
title: "Lec6"
author: "Me"
date: "2/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R4DS chapter 19,21

Practice turning the following code snippets into functions. Think about what each function does. What would you call it? How many arguments does it need? Can you rewrite it to be more expressive or less duplicative?
```{r}
mean(is.na(x))
x / sum(x, na.rm = TRUE)
sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)
```

```{r}
NA_frac<-function(x){
  mean(is.na(x))
}
NA_frac(c(1,2,3,NA))

get_fraction<-function(x){
  x / sum(x, na.rm = TRUE)
}
get_fraction(c(1,2,3,NA))

get_CV<-function(x){  #  coefficient of variation
  sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)
}
get_CV(c(1,2,3,NA))
```


Write both_na(), a function that takes two vectors of the same length and returns the number of positions that have an NA in both vectors.

```{r}
both_na<-function(a,b){
  if (length(a)!=length(b)){
    stop("Not same length",call. = TRUE)
  }
  sum(is.na(a)&is.na(b))
}
both_na(c(1,NA,2,NA),c(3,NA,12,444))
```

What do the following functions do? Why are they useful even though they are so short?
```{r}
is_directory <- function(x) file.info(x)$isdir   # whether x is a directory/folder
is_readable <- function(x) file.access(x, 4) == 0  # whether the file is readable
```

19.3.1 Exercises
Read the source code for each of the following three functions, puzzle out what they do, and then brainstorm better names.
```{r}
f1 <- function(string, prefix) {
  substr(string, 1, nchar(prefix)) == prefix
}
f1(c("abs","adidas"),"ab")  # check if the prefix is the first part of the string

f2 <- function(x) {
  if (length(x) <= 1) return(NULL)
  x[-length(x)]
}
f2(c(1,2,3,5))  # get rid of the last element

f3 <- function(x, y) {
  rep(y, length.out = length(x))
}
f3(c(1:10),c("a","b","c"))  # repeat vector y until get to the length of x
```

19.4.4 Exercises
Write a greeting function that says “good morning”, “good afternoon”, or “good evening”, depending on the time of day. (Hint: use a time argument that defaults to lubridate::now(). That will make it easier to test your function.)

```{r}
greeting<-function(){
  time<-lubridate::hour(lubridate::now())
  time
  if (time<12){
    print("Good morning")
  } else if(time<18){
    print("Good afternoon")
  } else{
    print("Good evening")
  }
}
greeting()
```


Implement a fizzbuzz function. It takes a single number as input. If the number is divisible by three, it returns “fizz”. If it’s divisible by five it returns “buzz”. If it’s divisible by three and five, it returns “fizzbuzz”. Otherwise, it returns the number. Make sure you first write working code before you create the function.
```{r}
fizzbuzz<-function(x){
  if (x%%5==0 & x%%3==0){
    print("fizzbuzz")
  } else if(x%%5==0){
    print("buzz")
  }else if(x%%3==0){
    print("fizz")
  }else{
    print(x)
  }
  
}
fizzbuzz(15)
fizzbuzz(1)
fizzbuzz(25)
fizzbuzz(18)
fizzbuzz(10123)

```

How could you use cut() to simplify this set of nested if-else statements?
```{r}
if (temp <= 0) {
  "freezing"
} else if (temp <= 10) {
  "cold"
} else if (temp <= 20) {
  "cool"
} else if (temp <= 30) {
  "warm"
} else {
  "hot"
}
```

```{r}
f1<-function(x){
  cut(x,c(-Inf,0,10,20,30,Inf),labels=c("freezing","cold","cool","warm","hot"))
}
f1(c(-100,0,4,15,276))
```

How would you change the call to cut() if I’d used < instead of <=? What is the other chief advantage of cut() for this problem? (Hint: what happens if you have many values in temp?)
```{r}
f1<-function(x){
  cut(x,c(-Inf,0,10,20,30,Inf),labels=c("freezing","cold","cool","warm","hot"),right=FALSE)
}
f1(c(-100,0,4,15,276))
```
What happens if you use switch() with numeric values?

What does this switch() call do? What happens if x is “e”?
```{r}
try<-function(x){
  switch(x, 
  a = ,     
  b = "ab",
  c = ,
  d = "cd"
)
}
# it will return the next value it meets
try("e")
try("a")
try("b")
```
Experiment, then carefully read the documentation.


21.2.1 Exercises
Write for loops to:
Compute the mean of every column in mtcars.
```{r}
mtcars
for (nm in names(mtcars)){
  print(mean(mtcars[[nm]],na.rm = TRUE))
}
```

Determine the type of each column in nycflights13::flights.
```{r}
library(nycflights13)
output<-vector("list",ncol(flights))
names(output)<-names(flights)
for (i in names(flights)) {
  output[[i]] <- class(flights[[i]])
}
output
```

Compute the number of unique values in each column of iris.
n_distinct(iris[[i]])
Generate 10 random normals from distributions with means of -10, 0, 10, and 100.
```{r}
mean_list<-c(-10,0,10,100)
for (i in seq_along(mean_list)){
  print(rnorm(10,mean=mean_list[i]))
}
```


