---
title: "Session_3_exercise_do_them_myself"
author: "Me"
date: "1/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

# chapter 8, chapter 5 (before 5.6)

5.2.4 Exercises
Find all flights that

Had an arrival delay of two or more hours

```{r}
filter(flights,arr_delay>120)
```

Flew to Houston (IAH or HOU)    Airport three letter code, FAA airport code.
```{r}
filter(flights,dest=="IAH"| dest=="HOU")
filter(flights,dest %in% c("IAH","HOU"))
```

Were operated by United, American, or Delta     Two letter abbreviation.
```{r}
filter(flights,carrier %in% c("UA","AA","DL"))
```
Departed in summer (July, August, and September)
```{r}
filter(flights,6<month & month<10)  # look at the dep_time, why the first one is "1"? it is 0:01 am.
filter(flights,between(month,7,9)) 
```

Arrived more than two hours late, but didn’t leave late

```{r}
filter(flights,arr_delay>120 & dep_delay<=0)
filter(flights,dep_delay<=0) # flights that didn't leave late
```

Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights,dep_delay>60 & dep_delay-arr_delay>30)  # save more than 30 min during the flight
```

Departed between midnight and 6am (inclusive)
Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?

```{r}
filter(flights,between(dep_time,0,600))  # Are the borders included in the results
# or is it >0 or >=0? could check the "month" for an answer
```


How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}
filter(flights,is.na(dep_time)) 
#  dep_time, dep_delay, arr_time, arr_delay, air_time. The f;ights were cancelled.
# is there any flight that has a dep_time but doesn't have a arr_time?
filter(flights,is.na(arr_time) & ! is.na(dep_time)) 
# Don't worry. No air crash in all flights that departed NYC (i.e. JFK, LGA or EWR) in 2013.
# Due to some reasons like weather, these flights didn't land at their original destinations.
```



Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)

```{r}
NA^0 # NA to the 0th power
NA|TRUE # anything else OR TRUE is always TRUE
NA & FALSE # anything else AND FALSE is always FALSE
NA * 0
Inf * 0  # infinite
# NaN (“Not a Number”) means 0/0
# NA (“Not Available”) is generally interpreted as a missing value
```




5.3.1 Exercises
How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).

```{r}
# need to specific a variable, like dep_delay
# put all rows with a missing value for dep_delay at the start of the table
head(arrange(flights, desc(is.na(dep_delay))))
arrange(flights, desc(is.na(dep_delay)))

```

Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
arrange(flights,desc(dep_delay))
arrange(flights,desc(arr_delay-dep_delay))  # the flights that delay the most during the flight
arrange(flights,desc(arr_delay),dep_time)
```

Sort flights to find the fastest (highest speed) flights.
```{r}
arrange(flights,desc(distance/air_time))
```

Which flights travelled the farthest? Which travelled the shortest?
```{r}
arrange(flights,desc(distance)) # to Honolulu
arrange(flights,distance) # the flight is from Newark to LaGuardia... # PHL Philadelphia
```


5.4.1 Exercises
Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.

```{r}
select(flights,dep_time, dep_delay, arr_time, arr_delay)
select(flights, starts_with('dep'), starts_with('arr'))
```


What happens if you include the name of a variable multiple times in a select() call?

```{r}
select(flights,dep_time,dep_time)  # only select once
```

What does the any_of() function do? Why might it be helpful in conjunction with this vector?
vars <- c("year", "month", "day", "dep_delay", "arr_delay")Copy
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights,any_of(vars))  # it select the variables in "vars"
select(flights,vars)
```


Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
select(flights, contains("TIME"))Copy
```{r}
select(flights, contains("TIME")) # The original code intends to select variables containing the characters TIME. However, the selected variables contain the lower case time instead. By default, contains() is not case sensitive.
select(flights,contains("TIME",ignore.case = FALSE)) # now the lower case time is not considered
```


5.5.2 Exercises
Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
flights <- mutate(flights,
                  dep_time_mins = dep_time %/% 100 * 60 + dep_time %% 100,   # Quotient and remainder
                  sched_dep_time_mins = sched_dep_time %/% 100 * 60 + sched_dep_time %% 100)

select(flights, starts_with('dep_time'), starts_with('sched'))
```


Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

```{r}
flights_new<-select(flights,air_time, arr_time, dep_time)
flights_new<-mutate(flights_new,
                    flight_time=arr_time - dep_time)
flights_new   #  the arr_time and dep_time are in the format of hour and minutes
flights_new<-mutate(flights_new,
                    real_flight_time=arr_time %/% 100 * 60 +arr_time %% 100 - dep_time%/% 100 * 60 -dep_time %% 100)
flights_new   # they are still different
```


Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?

```{r}
select(flights, dep_time, sched_dep_time, dep_delay) # dep_time and sched_dep_time are in clock format. Their difference in minutes is given by dep_delay.
```

Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().

```{r}
# not mentioned in the class?
head(arrange(flights, min_rank(desc(dep_delay))), 10)
# min_rank() is equivalent to rank() method with the argument ties.method = 'min. It assigns every tied element to the lowest rank.
```

What does 1:3 + 1:10 return? Why?

```{r}
1:3 + 1:10 #  the shorter vector with be ‘repeated’ to match the length of the longer vector.
```


What trigonometric functions does R provide?

```{r}
?sin
```














