---
title: "Session_4_exercise_do_them_myself"
author: "Me"
date: "2/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

# R4DS 5.6, 7.3, 7.7, 10, 11
```{r}
flights
```

5.6.7 Exercises
Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:

```{r}
# how many different flights are there? What are their min(arr_delay) like?
# select flights that are not cancelled
not_cancelled<-filter(flights,!is.na(dep_time) & !is.na(arr_time))
not_cancelled%>%
  group_by(flight)%>%
  summarise(
    min_delay=min(arr_delay),
    n=n()
    )#%>%
  # filter(is.na(min_delay))
```

A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.
Maybe this means two different categories
A flight is 15 minutes early 50% of the time
```{r}
# select flights that are not cancelled
not_cancelled<-filter(flights,!is.na(dep_time) & !is.na(arr_time))
not_cancelled%>%
  group_by(flight)%>%
  summarise(
    median_delay=median(arr_delay,na.rm = TRUE),  # deal with missing values!!!
    n=n()
    )%>%   # mutate generate something for each row, summarise generate one value for all the data in one group
  filter(median_delay<=-15) # these flights are 15min earlier for half of the chances
```
A flight is 15 minutes late 50% of the time
```{r}
# select flights that are not cancelled
not_cancelled<-filter(flights,!is.na(dep_time) & !is.na(arr_time))
not_cancelled%>%
  group_by(flight)%>%
  summarise(
    median_delay=median(arr_delay,na.rm = TRUE),  # deal with missing values!!!
    n=n()
    )%>%   # mutate generate something for each row, summarise generate one value for all the data in one group
  filter(median_delay>=15) # these flights are 15min late for half of the chances
```
A flight is always 10 minutes late.
```{r}
# select flights that are not cancelled
not_cancelled<-filter(flights,!is.na(dep_time) & !is.na(arr_time))
not_cancelled%>%
  group_by(flight)%>%
  summarise(
    min_delay=min(arr_delay,na.rm = TRUE),  # deal with missing values!!!
    n=n()
    )%>%   # mutate generate something for each row, summarise generate one value for all the data in one group
  filter(min_delay>=10)
```

A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.
The same idea as above

99% of the time a flight is on time. 1% of the time it’s 2 hours late.

Which is more important: arrival delay or departure delay?




Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).

```{r}
# what does wt mean?
?count()
```

```{r}
# select flights that are not cancelled
not_cancelled<-filter(flights,!is.na(dep_time) & !is.na(arr_time))
not_cancelled%>%
  group_by(dest)%>%
  summarise(
    n=n()
    )
```
```{r}
# select flights that are not cancelled
not_cancelled<-filter(flights,!is.na(dep_time) & !is.na(arr_time))
not_cancelled%>%
  group_by(tailnum)%>%
  summarise(
    sum_dis=sum(distance)
    )
```


Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?
dep_delay? As it never took off.

Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?

```{r}
flights%>%
  group_by(year,month,day)%>%
  summarise(
    n_cancel=sum(is.na(dep_delay)), # number of cancelled flights
    n=n(),
    average_dep_delay=mean(dep_delay,na.rm=TRUE),
    average_arr_delay=mean(arr_delay,na.rm=TRUE)
  )%>%
  ggplot()+
  geom_point(aes(x=n_cancel/n,y=average_dep_delay))
```


Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))
```{r}
not_cancelled<-filter(flights,!is.na(dep_time) & !is.na(arr_time))
not_cancelled%>%
  group_by(carrier)%>%
  summarise(
    ave_delay=mean(arr_delay,na.rm=TRUE)
    )%>%
  arrange(desc(ave_delay))
```

```{r}
not_cancelled<-filter(flights,!is.na(dep_time) & !is.na(arr_time))
not_cancelled%>%
  ggplot()+
  geom_boxplot(aes(x=carrier,y=arr_delay))
```

```{r}
flights %>%
  group_by(carrier, dest) %>% 
  summarise(
    n=n()
    )
```


What does the sort argument to count() do. When might you use it?

```{r}
not_cancelled<-filter(flights,!is.na(dep_time) & !is.na(arr_time))
not_cancelled%>%
  group_by(flight)%>%
  count(sort=TRUE)  # arrange data in the descending order
```




7.3.4 Exercises
Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

```{r}
diamonds
```

```{r}
ggplot(diamonds)+
  geom_histogram(aes(x=x),binwidth = 0.01)
```

```{r}
ggplot(diamonds)+
  geom_histogram(aes(x=y),binwidth = 0.01)+
  xlim(0,15)
```
```{r}
ggplot(diamonds)+
  geom_histogram(aes(x=z),binwidth = 0.01)+
  xlim(0,10)
```

Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)

```{r}
ggplot(diamonds)+
  geom_histogram(aes(x=price),binwidth = 10) # a range near 1500 is missing
```
```{r}
diamonds%>%
  filter(price<2000 , price>1000)%>%
ggplot()+
  geom_histogram(aes(x=price),binwidth = 10) # a range near 1500 is missing
# There are no diamonds with a price of $1,500
```

How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?
```{r}
diamonds%>%
  filter(carat==1)%>%  # 23 and 1558
  count()
```

Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?




10.5 Exercises

How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).
```{r}
is_tibble(diamonds)
is_tibble(flights)
is_tibble(mtcars)
```

Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?
```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```
If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?
df[[var]]
Practice referring to non-syntactic names in the following data frame by:

Extracting the variable called 1.

Plotting a scatterplot of 1 vs 2.

Creating a new column called 3 which is 2 divided by 1.

Renaming the columns to one, two and three.

annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)Copy
What does tibble::enframe() do? When might you use it?

What option controls how many additional column names are printed at the footer of a tibble?
