---
title: "Session 2"
author: "RSR, edited by me"
date: "1/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# 3.5.1
3. What plots does the following code make? What does . do?
Possible Answer: fomular. if you put a "." then you could draw panels with only one variable

```{r}
ggplot(data = mpg) +
 geom_point(mapping = aes(x = displ, y = hwy)) +
 facet_grid(drv ~ cyl) # each row is a different "drv", each column is a different "cyl"

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)  # each panel will occupy the whole row, each row is a different "drv"

ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)   # each panel will occupy the whole column
```

4. Take the first faceted plot in this section:
What are the advantages to using faceting instead of the colour aesthetic? What are the disadvantages? How might the balance change if you had a larger dataset?
Possible answer: A:If the researcher is color-blind? If there are too many points and they overlap with each other? The number of combinations could easily be seen. The relationship between thw two variables could be detected.
D: If two panels are very similar, it's easier to tell their difference if you put one over the other.

```{r}
ggplot(data = mpg) +
 geom_point(mapping = aes(x = displ, y = hwy,color=interaction(drv, cyl))) # deal with all combinations of two variables
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class,scales = "free")  # the x and y axis scales are different for each panel
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class,scales = "free_x") +  # the x and y axis scales are different for each panel
  ylim(0,NA)
```

```{r}
ggplot(data = mpg) + 
  geom_histogram(mapping = aes(x = hwy))+
  coord_flip()  # now the bars are horizontal
```


```{r}
ggplot(data = mpg) + 
  geom_histogram(mapping = aes(x = hwy),binwidth = 2)
```

```{r}
ggplot(data = mpg) + 
  geom_histogram(mapping = aes(x = hwy),bins = 50)
```

```{r}
ggplot(data = mpg) + 
  geom_histogram(mapping = aes(y = hwy))
```
```{r}
ggplot(data = mpg) + 
  geom_line(mapping = aes(x =displ, y = hwy))+
  geom_point(mapping = aes(x =displ, y = hwy))+
  geom_histogram(mapping = aes(x=displ),alpha=0.5)

```

# 3.6.1
2. Run this code in your head (copy/paste allowed) and predict what the output will look 
like. Then, run the code in R and check your predictions.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point(show.legend=F) + 
  geom_smooth(se = F,show.legend=F)
```

4. What does the se argument to geom_smooth() do?
PA:whether show the error bar (area?) or not
3. What does show.legend = FALSE (in the geom call, stick in in for problem 2) do? 
PA:It looks like it only works when you put it in the parentesis of the geom lines.
-> can put show.legend into the ggplot() call?  doesn't work, why not?
PA:No. Ask the developers. Haha.
5. Will these two graphs look different (copy/paste allowed)? Why/why not?

PA:They are the same.
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

6. This is a good one
Recreate the R code necessary to generate the following graphs (see the page)

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(size=3) + 
  geom_smooth(se = FALSE,size=3)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, group = drv)) + 
  geom_point(size=3) + 
  geom_smooth(se = FALSE,size=3)

```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```


```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(aes(color = drv)) + 
  geom_smooth(se = FALSE)
```


```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(aes(color = drv)) + 
  geom_smooth(aes(linetype = drv), se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point(color = "white", size = 4) +
  geom_point()
```

